---
title: "Wage_visualization"
author: "Team_10"
date: "3/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd('F:/GMU/STAT515/Mid Term Project')
```

```{r}
library(tidyverse)
library(ggthemes)
library(MASS)
source('D:/Spring_2021/STAT 515_Prof.ScottBruce/hw.R')
library(dplyr)
library(choroplethr)
library(choroplethrMaps)
library(RColorBrewer)
library(acs)

library(micromapST)
library(plotly)
```


```{r}
wage <- read.csv("D:/Spring_2021/STAT 515_Prof.ScottBruce/After final/miderm-final-projects/Minimum Wage.csv")
head(wage)
```
```{r}
stateabb <- read.csv("D:/Spring_2021/STAT 515_Prof.ScottBruce/After final/miderm-final-projects/stateabb.csv")
head(stateabb)
```

preprocessing for dynamic us map for effective min wage change accross years
```{r}

x_wage = wage%>%
  inner_join(stateabb , by.x = State ,by.y = State  )%>%
  select(Abbreviation,State, Year, min_wage = State.Minimum.Wage)%>%
  mutate(hover = paste(State , "\n$",min_wage))


```

dynamic us map for effective min wage change accross years

```{r}

plot_geo(x_wage,
        locationmode ='USA-states',
        frame = ~Year
          )%>%
  add_trace(locations = ~Abbreviation,
            z = ~min_wage,
            color = ~min_wage,
            text = ~hover,
            hoverinfo = 'text'
            ) %>%
  layout(geo = list(scope = 'usa'),
         title = 'min wage title'
         )%>%
  config(displayModeBar=FALSE)%>%
  colorbar(tickprefix = "$" )


```
micromap of all states in 1968 and 2020



```{r}

mic_wage = wage%>%
  filter(Year == 1968 )%>%
  dplyr :: select(State ,Effective.Minimum.Wage)

mic_wagex = wage%>%
  filter(Year == 2020 )%>%
  dplyr :: select(State ,Effective.Minimum.Wage)

m_wage = merge(mic_wage,mic_wagex,by='State')

names(m_wage)[names(m_wage) == "Effective.Minimum.Wage.x"] <- "ewg_68"
names(m_wage)[names(m_wage) == "Effective.Minimum.Wage.y"] <- "ewg_20"



panelDesc <- data.frame(
  type = c('map','id','bar','bar'),
  lab1 = c('' ,'','1968','2020'),
  lab3 = c('','','',''),
  col1 = c(NA,NA,'ewg_68','ewg_20')
)


fName = "mic.pdf"
pdf(file=fName,width=7.5,height=10)

micromapST( m_wage, panelDesc,
            rowNamesCol = 'State',
            rowNames = 'full',
            plotNames = 'full',  
            sortVar = 'ewg_20',ascend = FALSE,
            title = c("State Minimum Wage in USA by States",
                      "in 1968 , 2020"),
            ignoreNoMatches = TRUE)

dev.off()

```

state or fed preprocessing for chlororpeth
```{r}
wage$value = ifelse(wage$Effective.Minimum.Wage == wage$Federal.Minimum.Wage,"fed","state")
wage$region =  tolower(wage$State)
```
state or fed 1968
```{r}
sf_1968 = wage%>%
  filter(Year==1968)%>%
  dplyr :: select(region,value)

state_choropleth(sf_1968,
                 num_color = 4,
                 title = "US 1968 Minimum wage usage by States"
                 ) +
  theme(plot.title = element_text(hjust = .5) ) +
  scale_fill_manual(values = c("#DF6589FF",'#00B1D2FF'),name = "Minimum Wage Usage", labels= c('Federal','State'))
  

```

state or fed 2020

```{r}

sf_2020 = wage%>%
  filter(Year==2020)%>%
  dplyr :: select(region,value)


state_choropleth(sf_2020,
                 num_color = 4,
                 title = "US 2020 Minimum wage usage by States",
                 ) +
  theme(plot.title = element_text(hjust = .5) ) +
  scale_fill_manual(values = c("#DF6589FF",'#00B1D2FF'),name = "Minimum Wage Usage", labels= c('Federal','State'))

```
wages of fed to state changing states from 1968 - 2020 
```{r}

fos_wage = wage[1892:2863,]
fos_wage 

fos_wage = wage[1892:2863,]%>%
  filter(State %in% m_wage2$State)%>%
  select(Year,State , Effective.Minimum.Wage)

names(fos_wage)[names(fos_wage) == "Effective.Minimum.Wage"] <- "ewg"

t <- list(
  family = "Arial Black",
  face = 'bold',
  size = 11)
 plot_ly(fos_wage,frame = ~Year , x = ~ewg , y = ~State)%>%
         layout(title = list(text='States that Flippped from using Federal to State Minimum Wage in USA'),
           yaxis = list(title = 'States'),
         xaxis = list(title = 'Effective Minimum Wage '),
         legend = list(x = 0.80, y = 0.90),
         font=t)

```
preprocessing for top5
```{r}
top5_all =wage%>%
  group_by(Year)%>%
  arrange(desc(Effective.Minimum.Wage))%>%
  slice(1:5)%>%
  dplyr :: select(Year,State,Effective.Minimum.Wage)
```


top from 1968 to 1995

```{r}

state_95 = top5_all%>%
  filter(Year==1995)

state_95 =   as.vector(state_95$State)


top5_95 = wage[1:1350,] %>%
  filter(State %in% state_95)%>%
  group_by(Year)%>%
  arrange(desc(Effective.Minimum.Wage))%>%
  slice(1:5)%>%
  dplyr :: select(Year,State,Effective.Minimum.Wage)

ggplot(top5_95 , aes(x =Year   ,y=Effective.Minimum.Wage ,group=State, color = State ) )+
  geom_line(size=1.5,alpha = 0.6)+
  geom_point(size = 2.5)+
  labs(title = 'Trend in TOP 5 States of 1995 with Higest Minimum Wage' , x = 'Year' , y = 'Effective Minimum Wage' )+
  scale_y_discrete( limits = c(1,2,3,4,5))+
  theme(
    legend.position = 'bottom',
    plot.title = element_text( face='bold' , size =15 ),
    axis.title.x = element_text(face='bold', size =12),
    axis.title.y = element_text(face='bold', size =12),
    legend.text = element_text( face = 'bold',size =10 ),
    legend.title = element_text( face = 'bold',size =13 ),
    #changing faecet label text type and background color
    strip.text.x = element_text(
      size = 12, color = "white", face = "bold"
    ),
    strip.background = element_rect(
      color="white", fill="#f3a953", size=0.5, 
    )
  )

```
Top 5 from 1995 - 2020

```{r echo=FALSE}
state_20 = top5_all%>%
  filter(Year==2020)

state_20 =   as.vector(state_20$State)

top5_20 = wage[1512:2862,]%>%
  filter(State %in% state_20)%>%
  group_by(Year)%>%
  arrange(desc(Effective.Minimum.Wage))%>%
  slice(1:5)%>%
  dplyr :: select(Year,State,Effective.Minimum.Wage)

par(mfrow = c(1,2))
ggplot(top5_20,aes(x =Year   ,y=Effective.Minimum.Wage ,group=State, color = State) )+
  geom_line(size=1.5,alpha = 0.8)+
  geom_point(size = 2.5)+
  labs(title = 'Trend of TOP 5 States with Higest Minimum Wage in 2020')+
  theme_fivethirtyeight()

ggplot(top5_20 , aes(x =Year   ,y=Effective.Minimum.Wage ,group=State, color = State ) )+
  geom_line(size=1.5,alpha = 0.6)+
  geom_point(size = 2.5)+
  labs(title = 'Trend of TOP 5 States with Higest Minimum Wage in 1995', x = 'Year' , y = 'Effective Minimum Wage' )+
  hw+
  theme(
    legend.position = 'bottom',
    plot.title = element_text( face='bold' , size =15 ),
    axis.title.x = element_text(face='bold', size =12),
    axis.title.y = element_text(face='bold', size =12),
    legend.text = element_text( face = 'bold',size =10 ),
    legend.title = element_text( face = 'bold',size =13 ),
    #changing faecet label text type and background color
    strip.text.x = element_text(
      size = 12, color = "white", face = "bold"
    ),
    strip.background = element_rect(
      color="white", fill="#f3a953", size=0.5, 
    )
  )
```

```{r}

top5_95x = top5_95%>%
  mutate(diff = paste('1968 - 1995'))

top5_20x = top5_20%>%
  mutate(diff = paste('1995 - 2020'))

xa = data.frame(top5_95x , top5_20x )

```




```{r}
plot_ly(top5_20 , frame = ~Year ,x =~Effective.Minimum.Wage  , y = ~State ,type = 'scatter',linetype = ~Effective.Minimum.Wage ,mode = 'lines+markers',groups = ~State)
```


micromap of all states in 1968 and 2020 that flipped from fed to state 

```{r}

mic_wage2 = wage%>%
  filter(Year == 1968 & value=='fed')%>%
  dplyr :: select(State , State.Minimum.Wage ,Effective.Minimum.Wage)


mic_wagex2 = wage%>%
  filter(Year == 2020 & value == 'state')%>%
  dplyr :: select(State , Effective.Minimum.Wage)

m_wage2 = merge(mic_wage2, mic_wagex2 ,by= 'State')

names(m_wage2)[names(m_wage2) == "State.Minimum.Wage"] <- "swg_68"
names(m_wage2)[names(m_wage2) == "Effective.Minimum.Wage.x"] <- "ewg_68"
names(m_wage2)[names(m_wage2) == "Effective.Minimum.Wage.y"] <- "ewg_20"


panelDesc <- data.frame(
  type = c('map','id','bar','bar'),
  lab1 = c('' ,'','1968','2020'),
  lab3 = c('','','',''),
  col1 = c(NA,NA,'ewg_68','ewg_20')
)


fName = "mic2.pdf"
pdf(file=fName,width=7.5,height=10)

micromapST( m_wage2, panelDesc,
            rowNamesCol = 'State',
            rowNames = 'full',
            plotNames = 'full',  
            sortVar = 'ewg_20',ascend = FALSE,
            title = c("State Minimum wage that changed from Fed to State in USA by States",
                      "in 1968 , 2020"),
            ignoreNoMatches = TRUE,
            axisScale="sn")

dev.off()

```

preprocessing for dynamic us map for min wage type change accross years
```{r}


wage$value = ifelse(wage$Effective.Minimum.Wage == wage$Federal.Minimum.Wage,"fed","state")
y_wage = wage%>%
  inner_join(stateabb , by.x = State ,by.y = State  )%>%
  select(Abbreviation,State, Year, sof = value)%>%
  mutate(hover = paste(State))


```

dynamic us map for effective min wage change accross years

```{r}

plot_geo(y_wage,
         locationmode ='USA-states',
         frame = ~Year
            )%>%
  add_trace(locations = ~Abbreviation,
           # z = ~sof,
            color = ~sof,
            text = ~hover,
            hoverinfo = 'text'
  ) %>%
  layout(geo = list(scope = 'usa'),
         title = 'sof'
  ) %>%
  config(displayModeBar=TRUE)%>%
  colorbar(tickprefix = "$" )


```
wage os states from 1968 to 2020
```{r}

plot_ly(wage,frame = ~Year , x = ~Effective.Minimum.Wage , y = ~State )%>%
         layout(xaxis = list(title = 'States in USA'),
         yaxis = list(title = 'Effective Minimum Wage '),
         legend = list(x = 0.80, y = 0.90))

```





```{r}
mic_wage3 = wage%>%
  filter(Year == 1968 & value=='state')%>%
  select(State , State.Minimum.Wage ,Effective.Minimum.Wage)


mic_wagex3 = wage%>%
  filter(Year == 2020 & value == 'fed')%>%
  select(State , Effective.Minimum.Wage)

m_wage3 = merge(mic_wage3, mic_wagex3 ,by= 'State')

sof_wage = wage%>%
  filter(State %in% m_wage3$State)%>%
  select(Year,State , Effective.Minimum.Wage)

names(sof_wage)[names(sof_wage) == "Effective.Minimum.Wage"] <- "ewg"



plot_ly(sof_wage,frame = ~Year , x = ~ewg , y = ~State)

```


```{r}
top5s = wage%>%
  filter(Year %in% c(1968,1995,2020))%>%
  group_by(Year)%>%
  arrange(desc(Effective.Minimum.Wage))%>%
  slice(1:5)%>%
  select(Year,State,Effective.Minimum.Wage)%>%
  summarise(sum_5 = sum(Effective.Minimum.Wage))

all = wage%>%
  filter(Year %in% c(1968,1995,2020))%>%
  group_by(Year)%>%
  summarize(all_s = sum(Effective.Minimum.Wage))

th_s = merge(top5s,all,by='Year')

rest_s = wage%>%
  filter(Year %in% c(1968,1995,2020))%>%
  group_by(Year)%>%
  arrange(desc(Effective.Minimum.Wage))%>%
  slice(6:54)%>%
  summarize(rest_s = sum(Effective.Minimum.Wage))

th5_s = merge(top5s , rest_s,by='Year')
xa = merge(top5s , rest_s,by='Year')
xa = merge(xa , all, by = 'Year')

xa = xa %>%
  mutate(pe_top = sum_5/ all_s , pe_mid = )
th_5s = th5_s%>%
  gather(key=type_of_data,value=sum_value,-Year)

ggplot(th_5s,aes(x = Year , y=sum_value ,fill = type_of_data) )+
  geom_bar(stat = 'identity',position = position_dodge(18))+
    scale_fill_manual( labels= c('sum_5'='Top - 5 States','rest_s'='States Except Top-5'), values=c("#DF6589FF",'#00B1D2FF'))+
  geom_text(aes(group = type_of_data,label=paste(sum_value)),
            position = position_dodge(18) , vjust = -1, size = 5)+
  labs(y = "Sum of Minimum Wages of States",
       x="Years",
       title="Effect of Top 5 states with others",fill='Sum of Minimum Wages over States')+
  hw+
  theme(
    legend.position = 'top',
    plot.title = element_text( face='bold' , size =20 ),
    axis.title.x = element_text(face='bold', size =20),
    axis.title.y = element_text(face='bold', size =20),
    legend.text = element_text( face = 'bold',size =10 ),
    legend.title = element_text( face = 'bold',size =13 ),
    #changing faecet label text type and background color
    strip.text.x = element_text(
      size = 12, color = "white", face = "bold"
    ),
    strip.background = element_rect(
      color="white", fill="#f3a953", size=0.5, 
    )
  )

```

```{r}

#open on full screen to view plot 
bot5_68  = wage%>%
  filter(Year==1968)%>%
  arrange(State.Minimum.Wage)%>%
  slice(1:5)

cp = wage%>%
  filter(Year %in% c(1968,1995,2020) , State %in% bot5_68$State)%>%
  mutate(w4 = Effective.Minimum.Wage*80)%>%
  select(Year,State,w4, cp = CPI.Average)

cp1 = cp%>%
  gather(key = cac, value = val, -State , -Year)%>%
  arrange(State)


ggplot(cp1 , aes(x = State , y = val , fill=cac))+
  geom_bar(stat = 'identity', position = 'dodge')+
  labs(title = 'Bottom 5 States Effective Minimum Wage Vs CPI Average', x = 'States' , y = 'Dollars', subtitle = '(Monthly Basis)' , fill='')+
  facet_grid(.~Year)+
  scale_fill_manual(labels = c('cp'='CPI Average','w4'='Effective Minimum Wage'), values =  c("#DF6589FF",'#00B1D2FF'))+
  geom_text(aes(label = paste(val)), position = position_dodge(width = 1  ), vjust=0 , size =5 )+
  hw+
  theme(
    legend.position = 'top',
    plot.title = element_text( face='bold' , size =18 ),
    axis.title.x = element_text(face='bold', size =15),
    axis.title.y = element_text(face='bold', size =15),
    legend.text = element_text( face = 'bold',size =10 ),
    legend.title = element_blank(),
    #changing faecet label text type and background color
    strip.text.x = element_text(
      size = 12, color = "white", face = "bold"
    ),
    strip.background = element_rect(
      color="white", fill="#f3a953", size=0.5, 
    )
  )
  
```

